Spatial dynamics of complex urban systems within an evolutionary theory frame

### 2020-11-20 Docs/Papers/HandbookEntropy/chapter.tex

\documentclass[10pt]{article}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{Fig1.jpg}
	\caption{Schematic representation of the different processes included in the macroscopic co- evolution model for cities and transportation networks. Cities are characterized by their population (circle area) and are linked by a transportation network (dashed lines). City population grows endogenously (Gibrat model, first panel), through spatial interactions (second panel) and feedback of flows (third panel). Network links evolve according to the flow they carry (fourth panel).\label{fig:fig1}}	
\end{figure}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{Fig2.jpg}	
	\caption{Schematic representation of the different processes included in the innovation diffusion urban evolution model. Cities are characterized by their population (circle area) and their innovation shares or genome (color boxes). Innovations are diffused between cities with a spatial interaction model (second panel). City sizes are updated according to spatial interactions and attractivity linked to the share of the dominant innovation (third panel). Genomes are mutated with possibly the introduction of new innovations (fourth panel).\label{fig:fig2}}
\end{figure}

@@

\begin{thebibliography}{}

\bibitem[Aghion et~al., 1998]{aghion1998endogenous}
Aghion, P., Ljungqvist, L., Howitt, P., Howitt, P.~W., Brant-Collett, M.,
  Garc{\'\i}a-Pe{\~n}alosa, C., et~al. (1998).
\newblock {\em Endogenous growth theory}.
\newblock MIT press.

\bibitem[Audretsch and Lehmann, 2005]{audretsch2005does}
Audretsch, D.~B. and Lehmann, E.~E. (2005).
\newblock Does the knowledge spillover theory of entrepreneurship hold for
  regions?
\newblock {\em Research policy}, 34(8):1191--1202.

\bibitem[Banos, 2013]{banos2013pour}
Banos, A. (2013).
\newblock {\em Pour des pratiques de mod{\'e}lisation et de simulation
  lib{\'e}r{\'e}es en g{\'e}ographie et SHS}.
\newblock PhD thesis.

\bibitem[Batty, 2013]{batty2013new}
Batty, M. (2013).
\newblock {\em The new science of cities}.
\newblock MIT press.

\bibitem[Brenner and Schmid, 2014]{brenner2014urban}
Brenner, N. and Schmid, C. (2014).
\newblock The ‘urban age’in question.
\newblock {\em International journal of urban and regional research},
  38(3):731--755.

\bibitem[Caldarelli et~al., 2018]{caldarelli2018physics}
Caldarelli, G., Wolf, S., and Moreno, Y. (2018).
\newblock Physics of humans, physics for society.
\newblock {\em Nature Physics}, 14(9):870--870.

\bibitem[Callebaut, 2012]{callebaut2012scientific}
Callebaut, W. (2012).
\newblock Scientific perspectivism: A philosopher of science’s response to
  the challenge of big data biology.
\newblock {\em Studies in History and Philosophy of Science Part C: Studies in
  History and Philosophy of Biological and Biomedical Sciences}, 43(1):69--80.

\bibitem[Chavalarias et~al., 2009]{chavalarias2009french}
Chavalarias, D., Bourgine, P., Perrier, E., Amblard, F., Arlabosse, F., Auger,
  P., Baillon, J.-B., Barreteau, O., Baudot, P., Bouchaud, E., et~al. (2009).
\newblock French roadmap for complex systems 2008-2009.

\bibitem[Clarke and Gaydos, 1998]{clarke1998loose}
Clarke, K.~C. and Gaydos, L.~J. (1998).
\newblock Loose-coupling a cellular automaton model and gis: long-term urban
  growth prediction for san francisco and washington/baltimore.
\newblock {\em International journal of geographical information science},
  12(7):699--714.

\bibitem[Cottineau, 2014]{cottineau2014evolution}
Cottineau, C. (2014).
\newblock {\em L'{\'e}volution des villes dans l'espace post-sovi{\'e}tique.
  Observation et mod{\'e}lisations.}
\newblock PhD thesis, Universit{\'e} Paris 1 Panth{\'e}on-Sorbonne.

\bibitem[Cottineau, 2017]{cottineau2017metazipf}
Cottineau, C. (2017).
\newblock Metazipf. a dynamic meta-analysis of city size distributions.
\newblock {\em PloS one}, 12(8):e0183919.

\bibitem[Cottineau, 2020]{cottineau2020metametazipf}
Cottineau, C. (2020).
\newblock Metametazipf. what do analyses of city size distributions have in
  common?
\newblock {\em arXiv preprint arXiv:2008.11473}.

\bibitem[Cottineau et~al., 2015a]{cottineau2015growing}
Cottineau, C., Chapron, P., and Reuillon, R. (2015a).
\newblock Growing models from the bottom up. an evaluation-based incremental
  modelling method (ebimm) applied to the simulation of systems of cities.
\newblock {\em Journal of Artificial Societies and Social Simulation}, 18(4):9.

\bibitem[Cottineau et~al., 2015b]{cottineau2015modular}
Cottineau, C., Reuillon, R., Chapron, P., Rey-Coyrehourcq, S., and Pumain, D.
  (2015b).
\newblock A modular modelling framework for hypotheses testing in the
  simulation of urbanisation.
\newblock {\em Systems}, 3(4):348--377.

\bibitem[Cura et~al., 2017]{cura2017old}
Cura, R., Cottineau, C., Swerts, E., Antonio~Ignazzi, C., Bretagnolle, A.,
  Vacchiani-Marcuzzo, C., and Pumain, D. (2017).
\newblock The old and the new: qualifying city systems in the world with
  classical models and new data.
\newblock {\em Geographical Analysis}, 49(4):363--386.

\bibitem[Denis, 2020]{denis2020more}
Denis, E. (2020).
\newblock More urban constructions for whom? drivers of urban built-up
  expansion across the world from 1990 to 2015.
\newblock In {\em Theories and Models of Urbanization}, pages 235--258.
  Springer.

\bibitem[Dupuy, 1987]{dupuy1987vers}
Dupuy, G. (1987).
\newblock Vers une th{\'e}orie territoriale des r{\'e}seaux: une application au
  transport urbain.
\newblock In {\em Annales de g{\'e}ographie}, pages 658--679. JSTOR.

\bibitem[Durand-Dast{\`e}s, 1991]{durand1991particulier}
Durand-Dast{\`e}s, F. (1991).
\newblock Le particulier et le g{\'e}n{\'e}ral en g{\'e}ographie.
\newblock In {\em Actes du sixi{\`e}me colloque de didactique de l'histoire, de
  la g{\'e}ographie et des sciences sociales}, volume~6, pages 209--219.
  Institut national de recherche p{\'e}dagogique.

\bibitem[Egidi et~al., 2020]{egidi2020long}
Egidi, G., Salvati, L., and Vinci, S. (2020).
\newblock The long way to tipperary: City size and worldwide urban population
  trends, 1950-2030.
\newblock {\em Sustainable Cities and Society}, page 102148.

\bibitem[Favaro and Pumain, 2011]{favaro2011gibrat}
Favaro, J.-M. and Pumain, D. (2011).
\newblock Gibrat revisited: An urban growth model incorporating spatial
  interaction and innovation cycles.
\newblock {\em Geographical Analysis}, 43(3):261--286.

\bibitem[Gibrat, 1931]{gibrat1931inegalites}
Gibrat, R. (1931).
\newblock {\em Les In{\'e}galit{\'e}s {\'E}conomiques, Librairie du Recueil
  Sirey, Paris}.

\bibitem[Giere, 2010]{giere2010scientific}
Giere, R.~N. (2010).
\newblock {\em Scientific perspectivism}.
\newblock University of Chicago Press.

\bibitem[Grossetti, 2020]{grossetti2020matiere}
Grossetti, M. (2020).
\newblock Mati{\`e}re sociale-introduction: {\`a} la recherche d'une ontologie
  robuste.

\bibitem[Hagerstrand et~al., 1968]{hagerstrand1968innovation}
Hagerstrand, T. et~al. (1968).
\newblock {\em Innovation diffusion as a spatial process.}
\newblock Chicago, USA: Univ. Chicago Press.

\bibitem[Halvorson, 2012]{halvorson2012scientific}
Halvorson, H. (2012).
\newblock What scientific theories could not be.
\newblock {\em Philosophy of science}, 79(2):183--206.

\bibitem[Huutoniemi et~al., 2010]{huutoniemi2010analyzing}
Huutoniemi, K., Klein, J.~T., Bruun, H., and Hukkinen, J. (2010).
\newblock Analyzing interdisciplinarity: Typology and indicators.
\newblock {\em Research policy}, 39(1):79--88.

\bibitem[Jiang et~al., 2015]{jiang2015zipf}
Jiang, B., Yin, J., and Liu, Q. (2015).
\newblock Zipf’s law for all the natural cities around the world.
\newblock {\em International Journal of Geographical Information Science},
  29(3):498--522.

\bibitem[Krugman, 1996]{krugman1996confronting}
Krugman, P. (1996).
\newblock Confronting the mystery of urban hierarchy.
\newblock {\em Journal of the Japanese and International economies},
  10(4):399--418.

\bibitem[Lane et~al., 2009]{lane2009complexity}
Lane, D., Pumain, D., van~der Leeuw, S.~E., and West, G. (2009).
\newblock {\em Complexity perspectives in innovation and social change},
  volume~7.
\newblock Springer Science \& Business Media.

\bibitem[Leclaire et~al., 2019]{leclaire2019retour}
Leclaire, M., Rey-Coyrehourcq, S., Charton, B., Arduin, H., Chapron, P.,
  Ch{\'e}rel, G., Delay, E., Gaston, B., Lavall{\'e}e, F., Passerat-Palmbach,
  J., et~al. (2019).
\newblock Retour d'exp{\'e}rience de l'{\'e}cole openmole" exmodelo",
  organis{\'e}e en partenariat avec le m{\'e}so-centre du criann.
\newblock In {\em Journ{\'e}es Calcul et Donn{\'e}es 2019}.

\bibitem[Lobo et~al., 2020]{lobo2020urban}
Lobo, J., Alberti, M., Allen-Dumas, M., Arcaute, E., Barthelemy, M.,
  Bojorquez~Tapia, L.~A., Brail, S., Bettencourt, L., Beukes, A., Chen, W.-Q.,
  et~al. (2020).
\newblock Urban science: Integrated theory from the first cities to sustainable
  metropolises.

\bibitem[Mesoudi, 2017]{mesoudi2017pursuing}
Mesoudi, A. (2017).
\newblock Pursuing darwin’s curious parallel: Prospects for a science of
  cultural evolution.
\newblock {\em Proceedings of the National Academy of Sciences},
  114(30):7853--7860.

\bibitem[Modica et~al., 2017]{modica2017methodological}
Modica, M., Reggiani, A., and Nijkamp, P. (2017).
\newblock Methodological advances in gibrat’s and zipf’s laws: a
  comparative empirical study on the evolution of urban systems.
\newblock In {\em Socioeconomic Environmental Policies and Evaluations in
  Regional Science}, pages 37--59. Springer.

\bibitem[Mustafa et~al., 2017]{mustafa2017coupling}
Mustafa, A., Cools, M., Saadi, I., and Teller, J. (2017).
\newblock Coupling agent-based, cellular automata and logistic regression into
  a hybrid urban expansion model (huem).
\newblock {\em Land Use Policy}, 69:529--540.

\bibitem[Openshaw, 1993]{openshaw1993modelling}
Openshaw, S. (1993).
\newblock Modelling spatial interaction using a neural net.
\newblock In {\em Geographic information systems, spatial modelling and policy
  evaluation}, pages 147--164. Springer.

\bibitem[Paulus, 2004]{paulus2004coevolution}
Paulus, F. (2004).
\newblock {\em Co{\'e}volution dans les syst{\`e}mes de villes: croissance et
  sp{\'e}cialisation des aires urbaines fran{\c{c}}aises de 1950 {\`a} 2000}.
\newblock PhD thesis, Universit{\'e} Panth{\'e}on-Sorbonne-Paris I.

\bibitem[Pumain, 1997]{pumain1997pour}
Pumain, D. (1997).
\newblock Pour une th{\'e}orie {\'e}volutive des villes.
\newblock {\em L'Espace g{\'e}ographique}, pages 119--134.

\bibitem[Pumain, 1998]{pumain1998urban}
Pumain, D. (1998).
\newblock Urban research and complexity.
\newblock In {\em The city and its sciences}, pages 323--361. Springer.

\bibitem[Pumain, 2003]{pumain2003approche}
Pumain, D. (2003).
\newblock Une approche de la complexit{\'e} en g{\'e}ographie.
\newblock {\em G{\'e}ocarrefour}, 78(1):25--31.

\bibitem[Pumain, 2005]{pumain2005cumulativite}
Pumain, D. (2005).
\newblock Cumulativit{\'e} des connaissances.
\newblock {\em Revue europ{\'e}enne des sciences sociales. European Journal of
  Social Sciences}, (XLIII-131):5--12.

\bibitem[Pumain, 2012]{pumain2012urban}
Pumain, D. (2012).
\newblock Urban systems dynamics, urban growth and scaling laws: The question
  of ergodicity.
\newblock In {\em Complexity theories of cities have come of age}, pages
  91--103. Springer.

\bibitem[Pumain, 2020]{pumain2020theories}
Pumain, D. (2020).
\newblock {\em Theories and Models of Urbanization}.
\newblock Springer.

\bibitem[Pumain et~al., 2012]{pumain2012theorie}
Pumain, D. et~al. (2012).
\newblock Une th{\'e}orie g{\'e}ographique pour la loi de zipf.
\newblock {\em R{\'e}gion et d{\'e}veloppement}, 36(31-54):125--150.

\bibitem[Pumain and Raimbault, 2020]{pumain2020conclusion}
Pumain, D. and Raimbault, J. (2020).
\newblock Conclusion: Perspectives on urban theories.
\newblock In {\em Theories and Models of Urbanization}, pages 303--330.
  Springer.

\bibitem[Pumain and Reuillon, 2017]{pumain2017urban}
Pumain, D. and Reuillon, R. (2017).
\newblock {\em Urban dynamics and simulation models}.
\newblock Springer.

\bibitem[Pumain and Robic, 1996]{pumain1996theoriser}
Pumain, D. and Robic, M. (1996).
\newblock Th{\'e}oriser la ville in derycke ph, huriot jm, pumain d.
\newblock {\em Penser la Ville. Th{\'e}ories et mod{\`e}les}.

\bibitem[Pumain and Sanders, 2013]{pumain2013theoretical}
Pumain, D. and Sanders, L. (2013).
\newblock Theoretical principles in interurban simulation models: a comparison.
\newblock {\em Environment and Planning A}, 45(9):2243--2260.

\bibitem[Pumain et~al., 2015]{pumain2015multilevel}
Pumain, D., Swerts, E., Cottineau, C., Vacchiani-Marcuzzo, C., Ignazzi, C.~A.,
  Bretagnolle, A., Delisle, F., Cura, R., Lizzi, L., and Baffi, S. (2015).
\newblock Multilevel comparison of large urban systems.
\newblock {\em Cybergeo: European Journal of Geography}.

\bibitem[Raimbault, 2017]{raimbault2017applied}
Raimbault, J. (2017).
\newblock An applied knowledge framework to study complex systems.
\newblock In {\em Complex Systems Design \& Management}, pages 31--45.

\bibitem[Raimbault, 2018]{raimbault2018calibration}
Raimbault, J. (2018).
\newblock Calibration of a density-based model of urban morphogenesis.
\newblock {\em PloS one}, 13(9):e0203516.

\bibitem[Raimbault, 2020a]{raimbault2020hierarchy}
Raimbault, J. (2020a).
\newblock Hierarchy and co-evolution processes in urban systems.
\newblock {\em arXiv preprint arXiv:2001.11989}.

\bibitem[Raimbault, 2020b]{raimbault2020indirect}
Raimbault, J. (2020b).
\newblock Indirect evidence of network effects in a system of cities.
\newblock {\em Environment and Planning B: Urban Analytics and City Science},
  47(1):138--155.

\bibitem[Raimbault, 2020c]{raimbault2020model}
Raimbault, J. (2020c).
\newblock A model of urban evolution based on innovation diffusion.
\newblock {\em Artificial Life Conference Proceedings}, (32):500--508.

\bibitem[Raimbault, 2020d]{raimbault2020modeling}
Raimbault, J. (2020d).
\newblock Modeling the co-evolution of cities and networks.
\newblock In {\em Handbook of Cities and Networks, Rozenblat C., Niel Z., eds.
  Edward Elgar Publishing. In press}.

\bibitem[Raimbault, 2020e]{raimbault2020relating}
Raimbault, J. (2020e).
\newblock Relating complexities for the reflexive study of complex systems.
\newblock In {\em Theories and models of urbanization}, pages 27--41. Springer.

\bibitem[Raimbault, 2020f]{raimbault2020unveiling}
Raimbault, J. (2020f).
\newblock Unveiling co-evolutionary patterns in systems of cities: a systematic
  exploration of the simpopnet model.
\newblock In {\em Theories and Models of Urbanization}, pages 261--278.
  Springer.

\bibitem[Raimbault et~al., 2019]{raimbault2019space}
Raimbault, J., Cottineau, C., Le~Texier, M., Le~Nechet, F., and Reuillon, R.
  (2019).
\newblock Space matters: Extending sensitivity analysis to initial spatial
  conditions in geosimulation models.
\newblock {\em Journal of Artificial Societies and Social Simulation}, 22(4).

\bibitem[Raimbault et~al., 2020a]{raimbault2020empowering}
Raimbault, J., Denis, E., and Pumain, D. (2020a).
\newblock Empowering urban governance through urban science: Multi-scale
  dynamics of urban systems worldwide.
\newblock {\em Sustainability}, 12(15):5954.

\bibitem[Raimbault et~al., 2020b]{raimbault2020scala}
Raimbault, J., Perret, J., and Reuillon, R. (2020b).
\newblock A scala library for spatial sensitivity analysis.
\newblock In {\em GISRUK 2020 Proceedings. Birkbeck, University of London}.

\bibitem[Raimbault and Pumain, 2019]{raimbault2019methods}
Raimbault, J. and Pumain, D. (2019).
\newblock Methods for exploring simulation models.
\newblock {\em Geographical Modeling: Cities and Territories}, 2:125--150.

\bibitem[Reggiani and Nijkamp, 2009]{reggiani2009complexity}
Reggiani, A. and Nijkamp, P. (2009).
\newblock {\em Complexity and spatial networks: in search of simplicity}.
\newblock Springer Science \& Business Media.

\bibitem[Reuillon et~al., 2015]{reuillon2015new}
Reuillon, R., Schmitt, C., De~Aldama, R., and Mouret, J.-B. (2015).
\newblock A new method to evaluate simulation models: the calibration profile
  (cp) algorithm.
\newblock {\em Journal of Artificial Societies and Social Simulation},
  18(1):12.

\bibitem[Rey-Coyrehourcq, 2015]{rey2015plateforme}
Rey-Coyrehourcq, S. (2015).
\newblock {\em Une plateforme int{\'e}gr{\'e}e pour la construction et
  l'{\'e}valuation de mod{\`e}les de simulation en g{\'e}ographie.(An
  integrated platform for building and evaluating model of simulation in
  geography).}
\newblock PhD thesis, Pantheon-Sorbonne University, Paris, France.

\bibitem[Rey-Coyrehourcq et~al., 2019]{rey2019calcul}
Rey-Coyrehourcq, S., Banos, A., and Raimbault, J. (2019).
\newblock Le calcul intensif en g{\'e}ographie: une tradition bien ancr{\'e}e.
\newblock In {\em JCAD Journ{\'e}es Calcul et Donn{\'e}es 2019}.

\bibitem[Robinson, 2016]{robinson2016comparative}
Robinson, J. (2016).
\newblock Comparative urbanism: new geographies and cultures of theorizing the
  urban.
\newblock {\em International Journal of Urban and Regional Research},
  40(1):187--199.

\bibitem[Robson, 1973]{robson1973urban}
Robson, B. (1973).
\newblock {\em Urban Growth: an approach}.
\newblock Methuen \& Co Ltd., London.

\bibitem[Roth, 2006]{roth2006reconstruction}
Roth, C. (2006).
\newblock Reconstruction failures: Questioning level design.
\newblock In {\em International Workshop on Epistemological Aspects of Computer
  Simulation in the Social Sciences}, pages 89--98. Springer.

\bibitem[Rozenblat, 2020]{rozenblat2020extending}
Rozenblat, C. (2020).
\newblock Extending the concept of city for delineating large urban regions
  (lur) for the cities of the world.
\newblock {\em Cybergeo: European Journal of Geography}.

\bibitem[Schmitt, 2014]{schmitt2014modelisation}
Schmitt, C. (2014).
\newblock {\em Mod{\'e}lisation de la dynamique des syst{\`e}mes de peuplement:
  de SimpopLocal {\`a} SimpopNet}.
\newblock PhD thesis, Universit{\'e} Panth{\'e}on-Sorbonne-Paris I.

\bibitem[Schmitt et~al., 2015]{schmitt2015half}
Schmitt, C., Rey-Coyrehourcq, S., Reuillon, R., and Pumain, D. (2015).
\newblock Half a billion simulations: Evolutionary algorithms and distributed
  computing for calibrating the simpoplocal geographical model.
\newblock {\em Environment and Planning B: Planning and Design},
  42(2):300--315.

\bibitem[Scott and Storper, 2015]{scott2015nature}
Scott, A.~J. and Storper, M. (2015).
\newblock The nature of cities: The scope and limits of urban theory.
\newblock {\em International journal of urban and regional research},
  39(1):1--15.

\bibitem[Suppe, 2000]{suppe2000understanding}
Suppe, F. (2000).
\newblock Understanding scientific theories: An assessment of developments,
  1969-1998.
\newblock {\em Philosophy of science}, 67:S102--S115.

\bibitem[Swerts et~al., 2018]{swerts2018diffuse}
Swerts, E., Denis, E., and Mukhopadhyay, P. (2018).
\newblock Diffuse urbanization and mega-urban regions in india: Between
  reluctant and restrictive urbanism?
\newblock In {\em International and Transnational Perspectives on Urban
  Systems}, pages 237--262. Springer.

\bibitem[Varenne, 2018]{varenne2018theories}
Varenne, F. (2018).
\newblock {\em Th{\'e}ories et mod{\`e}les en sciences humaines: le cas de la
  g{\'e}ographie}.
\newblock {\'E}ditions Mat{\'e}riologiques.

\bibitem[Winther, 2016]{winther2016structure}
Winther, R.~G. (2016).
\newblock The structure of scientific theories.
\newblock In {\em The Stanford Encyclopedia of Philosophy Edward N. Zalta
  (ed.)}.

\bibitem[Wu, 2020]{wu2020emerging}
Wu, F. (2020).
\newblock Emerging cities and urban theories: a chinese perspective.
\newblock In {\em Theories and models of urbanization}, pages 171--182.
  Springer.

\end{thebibliography}

>>>>

\documentclass[10pt,letterpaper]{article}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figures/Fig1.jpg}
	\caption{Schematic representation of the different processes included in the macroscopic co- evolution model for cities and transportation networks. Cities are characterized by their population (circle area) and are linked by a transportation network (dashed lines). City population grows endogenously (Gibrat model, first panel), through spatial interactions (second panel) and feedback of flows (third panel). Network links evolve according to the flow they carry (fourth panel).\label{fig:fig1}}	
\end{figure}

@@

\begin{figure}
	\includegraphics[width=\linewidth]{figures/Fig2.jpg}	
	\caption{Schematic representation of the different processes included in the innovation diffusion urban evolution model. Cities are characterized by their population (circle area) and their innovation shares or genome (color boxes). Innovations are diffused between cities with a spatial interaction model (second panel). City sizes are updated according to spatial interactions and attractivity linked to the share of the dominant innovation (third panel). Genomes are mutated with possibly the introduction of new innovations (fourth panel).\label{fig:fig2}}
\end{figure}

@@

\bibliographystyle{apalike}
\bibliography{biblio}
